Intentional release of native species undermines ecological stability

### 2022-12-24 rmd/manuscript.Rmd
edit: manuscript

The intentional release, however, can have far-reaching consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has stimulated a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019; @cucheroussetAreDomesticatedFreshwater2020]. Yet, current debates overlook the fact that we lack a quantitative assessment of the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.

>>>>

The intentional release, however, can have wide-ranging consequences than previously thought, as the impacts can propagate through a diversity of ecological interactions [@careyNativeInvadersChallenges2012]. This recognition has sparked a discussion of how the massive introduction of native species alters short-term ecosystem dynamics [@hasegawaAreHatcheryChum2018; @nelsonEcologicalImplicationsChanging2019; @cucheroussetAreDomesticatedFreshwater2020]. Yet, current debates overlook the fact that we rarely assessed the community-wide impact in the long-term. Ecological theory suggests that the stable coexistence of competing species through density-dependent feedback underpins the emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Intentional release may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@laikreCompromisingGeneticDiversity2010; @careyNativeInvadersChallenges2012; @kitadaEconomicEcologicalGenetic2018]. Hence, this form of species management may intervene in the ecological process that allows competing species to coexist, ultimately degrading the long-term community stability. Evidence for this hypothesis is lacking, however.
